---
title: "Replication Paper"
author: "Xiao Sun"
date: "10/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(repr)
# Change plot size to 16 x 9
options(repr.plot.width=16, repr.plot.height=9)
```

# Title: Women and economic statecraft: The negative impact international economic sanctions visit on women

#Notes: research 1971-2005, but dataset is 1970-2004. It is one year lagged.


```{r echo=FALSE, message=FALSE, warning=FALSE}
## load pkgs and data
library(dplyr)
library(MASS) # for logit, help("polr")
library(VGAM) # vglm
library(ggplot2)
library(nnet)
library(lmtest)
library(stargazer)
library(texreg)
library(xtable)
library(haven) # for read_dta
data <- read_dta("EJIR Full data.dta")
str(data)
names(data)
glimpse(data)
```

The authors mentions that three CIRI variables: women's economic rights (wecon), women's social rights (wosoc), and women's political rights (wopol) are temporally dependent. So they follow Hafner-Burton’s lead and claim that they correct for the autocorrelation in the CIRI models by creating binary variables for each of the four categories of the women’s rights variables. But they do not use these binary variables as DVs. Instead, they still use original ordinal variables. So I will also use the original variables and then reorder them.

## reorder ordinal DVs
```{r echo=FALSE, message=FALSE, warning=FALSE}
data$wecon <- factor(data$wecon, levels = c("0", "1", "2", "3"), ordered = T)
data$wosoc <- factor(data$wosoc, levels = c("0", "1", "2", "3"), ordered = T)
data$wopol <- factor(data$wopol, levels = c("0", "1", "2", "3"), ordered = T)
# check
str(data$wecon)
str(data$wosoc)
str(data$wopol)
```

## Create lagged variables
```{r echo=FALSE, message=FALSE, warning=FALSE}

str(data$lsanction_hse_dur) ## NULL, the authors did not include this variable in this dataset

# create lag sanction_hse_dur
#library("dplyr")
data <- data %>%                            # Add lagged column, lag sanction_hse_dur
  group_by(ccode) %>%
  mutate( lsanction_hse_dur= lag(sanction_hse_dur, n = 1, default = NA)) %>% 
  as.data.frame()

str(data$lsanction_hse_dur)

# compare the original variable and the lagged variable
str(data$sanction_hse_dur)
str(data$lsanction_hse_dur) # the first is NA 
```

The authors use a one-year lag (t − 1) of the time-variant variables to mitigate any simultaneity issues between the dependent and independent variables. Specifically, lagged-IVs allow they to make sure that the independent variables temporally precede the dependent variables and thus eliminate any incorrect direction of inference.

```{r echo=FALSE, message=FALSE, warning=FALSE}

## convert integers 0 and 1 to dummy variables/ nominal variables
### check the attributes
str(data$lsanction_hse) ## lagged sanction variable
str(data$lagwar) ##  Interstate War
str(data$lagcwar) ## Civil War
str(data$multilateral)## Multilateral Sanction 
str(data$laghumansanction)## Humanitarian Sanction 
str(data$lagpolity2) ## nominal political scores, democracy -10~10

## binary variables for 3 levels of women's economic rights
#data$lwecon1 <- factor(data$lwecon1, levels = c("0", "1"))
#data$lwecon2 <- factor(data$lwecon2, levels = c("0", "1"))
#data$lwecon3 <- factor(data$lwecon3, levels = c("0", "1"))

## binary variables for 3 levels of women's social rights
#data$lwosoc1 <- factor(data$lwosoc1, levels = c("0", "1"))
#data$lwosoc2 <- factor(data$lwosoc2, levels = c("0", "1"))
#data$lwosoc3 <- factor(data$lwosoc3, levels = c("0", "1"))

## binary variables for 3 levels of women's political rights
#data$lwopol1 <- factor(data$lwopol1, levels = c("0", "1"))
#data$lwopol2 <- factor(data$lwopol2, levels = c("0", "1"))
#data$lwopol3 <- factor(data$lwopol3, levels = c("0", "1"))

#data$lsanction_hse <- factor(data$lsanction_hse, levels = c("0", "1"))
#data$lagwar <- factor(data$lagwar, levels = c("0", "1"))
#data$lagcwar <- factor(data$lagcwar, levels = c("0", "1"))

data$wecon <- factor(data$wecon, levels = c("0", "1", "2", "3"), ordered = T)
data$wopol <- factor(data$wopol, levels = c("0", "1", "2", "3"), ordered = T)
data$wosoc <- factor(data$wosoc, levels = c("0", "1", "2", "3"), ordered = T)

## Areas
data$asia <- factor(data$asia, levels = c("0", "1"))
data$lamerica <-  factor(data$lamerica, levels = c("0", "1"))
data$ssafrica <- factor(data$ssafrica, levels = c("0", "1"))
data$nafrme <- factor(data$nafrme, levels = c("0", "1"))
```

## Explore data
```{r}
# 3 DVs, plots
#jpeg(file="dvplote.jpeg")
plot(data$wecon, main="women's eocnomic rights", col = "darkred")
#dev.off()

#jpeg(file="dvplotp.jpeg")
plot(data$wopol, main="women's political rights", col = "darkgreen")
#dev.off()

#jpeg(file="dvplots.jpeg")
plot(data$wosoc, main="women's social rights", col = "darkblue")
#dev.off()

se <- summary(data$wecon) 
ss <- summary(data$wosoc)
sp <- summary(data$wopol)


stargazer(se, ss, sp, type = "text")
```

```{r}
#IVs
## economic sanction

png(file="sanctionplot.png")
summary(data$lsanction_hse)
plot(data$lsanction_hse, type = "p", lwd = 0.5,
     main="Sanction Plot",
     xlab="A country in a year",
     ylab="Sanction (dummy)") # most data are 0s
dev.off()

png(file="sanctioncost.png")
plot(data$lagsanctioncost, type = "p", lwd = 0.5,
     main="Sanction Cost Plot",
     xlab="A country in a year",
     ylab="Sanction Cost")
dev.off()

png(file="logGDP.png")
plot(data$laggdplog, type = "p", lwd = 0.5,
     main="Countries' GDP",
     xlab="A country in a year",
     ylab="Log of GDP")
dev.off()

png(file="polity.png")
plot(data$lagpolity2, type = "p", lwd = 0.5,
     main="Polity Scores",
     xlab="A country in a year",
     ylab="Political Score")
dev.off()

```



# Creale Table 1  
## My replication results are basically the same as the original results

## Women's eoconomic status
```{r warning=FALSE}

## model 1
modeco1 <- polr(wecon ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwecon1 + lwecon2 + lwecon3,  method = "logistic", data=data, Hess=T)

# if delete " + lwecon1 + lwecon2 + lwecon3", is different.

summary(modeco1)
stargazer(modeco1,type='text') 

## model 2: Duration, Duration*GDP Human Sanctions, and Multilateral sanctions
modeco2 <- polr(wecon ~ laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwecon1 + lwecon2 + lwecon3, method ="logistic", data=data, Hess=T)

summary(modeco2)
stargazer(modeco2,type='text') ## the results are basically the same

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
## model 3

### The authors use xtreg () in Stata for model 3 for fixed effects, so I will use plm() in R

## install.packages("plm" )
library("plm" )

str(data$fem_labor_first) ## initial female labor score
# fem.1 <- select(data1, 'fem_labor_first')
# str(data1$wdifemlabor) ## The authors use DV d.wdifemlabor, but I only see wdifemlabor, I guess it is about differencing this variable.

## They mentioned unit root issues: "diagnostic tests revealed that the Female Labor Participation variable has a strong unit root process. In order to correct for the lack of stationarity in this variable, we first-difference the data by subtracting the past value of the variable from the current value (Yt – Yt-1)."

## create a new variable: d.wdifemlabor

data <- data %>%                            # Add lagged column, lag wdifemlabor
  group_by(ccode) %>%
  dplyr::mutate(lwdifemlabor= dplyr::lag(wdifemlabor, n = 1, default = NA))

data <- data %>%
             mutate(d.wdifemlabor = wdifemlabor - lwdifemlabor)  # problem, this is calculated by 1970-2004 data minus 1969-2003 data. But instead it should be 1971-2005 - 1970-2004. The author did not provide me a complete dataset.

data$wdifemlabor
data$d.wdifemlabor

modeco3 <- plm (d.wdifemlabor ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + fem_labor_first, data = data, index = c( "ccode", "year")) # I set index for state and year, but unsure if it's ok.
summary(modeco3)
stargazer(modeco3,type='text') # the result is different from the original table.

## Model 4: Duration, Duration*GDP, Human Sanctions, and Multilateral sanctions
modeco4 <-  plm(d.wdifemlabor ~ laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + fem_labor_first, data = data, index = c("ccode", "year")) # no area displays in the table and results are different.

summary(modeco4)
stargazer(modeco4,type='text')
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# All tables in Table 1
stargazer(modeco1, modeco2,modeco3, modeco4,type='latex', digits = 3, no.space = TRUE,
          column.sep.width = "1pt", font.size = "small",
          title = "The effect of sanctions on women’s economic conditions",
          dep.var.labels = c("Women's Economic Rights", "Female Labor Participation"),
          covariate.labels = c("Economic Sanctions", "Sanction*GDP", "GDP Per Capita", "Sanction Duration", "Sanction Duration*GDP", "Sanction Costs", "Multilateral Sanction", "Humanitarian Sanction", "Democracy", "Economic Openess", "War", "Civil War", "Peace Years", "Asia/Pacific", "Latin America/Caribbean", "Sub-Saharan Africa", "M.East/N.Africa", "Initial Female Labor")) 
```


# Create Table 2: WOMEN'S SOCIAL AND POLITICAL STATUS
```{r, warning=FALSE}
## women's social rights
## model 1
modso1 <- polr(wosoc ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwosoc1 + lwosoc2 + lwosoc3, data=data, Hess=T)
summary(modso1)
stargazer(modso1,type='text') ## the results are basically the same

## model 2: Duration, Duration*GDP, Human Sanctions, and Multilateral sanctions
modso2 <- polr(wosoc ~ laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwosoc1 + lwosoc2 + lwosoc3, data=data, Hess=T)

summary(modso2)
stargazer(modso2,type='text') ## the results are basically the same
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
## women's political rights
## model 1
modpo1 <- polr(wopol ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwopol1 + lwopol2 + lwopol3, data=data, Hess=T) 

summary(modpo1)
stargazer(modpo1,type='text') ## the results are basically the same

## model 2
modpo2 <- polr(wopol ~ laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwopol1 + lwopol2 + lwopol3, data=data, Hess=T)

summary(modpo2)
stargazer(modpo2,type='text') ## the results are basically the same

# All tables in Table 2

stargazer(modso1, modso2,modpo1, modpo2, type='latex', digits = 3,no.space = TRUE,
          column.sep.width = "1pt", font.size = "small",
          title = "The effect of sanctions on women’s social and political conditions",
          dep.var.labels = c("Women's Social Rights", "Women’s Political Rights"),
          covariate.labels = c("Economic Sanctions", "Sanction*GDP", "GDP Per Capita", "Sanction Duration", "Sanction Duration*GDP", "Sanction Costs", "Multilateral Sanction", "Humanitarian Sanction", "Democracy", "Economic Openess", "War", "Civil War", "Peace Years", "Asia/Pacific", "Latin America/Caribbean", "Sub-Saharan Africa", "M.East/N.Africa")) 
```

# PRA test #
```{r}
library(brant)
```

### PRA test for model 1 and 2, women's economic rights
```{r, warning=FALSE}

# model 1 PRA test
sink("modeco1pra.txt")
brant(modeco1, by.var = T) # Omnibus, lsanction_hse, lagsanctiongdp, laggdplog, peace, asia, ssafrica, and nafrme violate PRA.
sink()

# model 2  PRA test
sink("modeco2pra.txt")
brant(modeco2, by.var = T) # Omnibus, laggdplog, laghumansanction, peace, asia,and ssafrica violate PRA.
sink()
```

### PRA test for model 1 and 2, women's social rights
```{r}
# model 1 PRA test
sink("modso1pra.txt")
#brant(modso1, by.var = T) # Omnibus, laggdplog, lagpolity2, asia, ssafrica, and nafrme violate PRA.
sink()

# model 2  PRA test
sink("modso2pra.txt")
#brant(modso2, by.var = T) # Omnibus, laggdplog, lagpolity2, asia, ssafrica, and nafrme violate PRA.
sink()
```

### PRA test for model 1 and 2, women's political rights. 
```{r, warning=FALSE}
# model 1 PRA test
sink("modpo1pra.txt")
library(brant)
brant(modpo1, by.var = T) # Omnibus, laggdplog, lagpolity2, asia, ssafrica, nafrme, lwosoc1, 2,3 violate PRA
sink()

# model 2  PRA test
sink("modpo2pra.txt")
#brant(modpo2, by.var = T) # Omnibus, laggdplog, lagpolity2, asia, ssafrica, nafrme, lwosoc1, 2,3 violate PRA
sink()
```


##dataset extension

### note: try different variables : fertility, literacy_fem, unemploy_fem, agri_fem, indus_fem, serv_fem, vulemp_fem, seceducattained, terteducattained, spendedc. Only fertility and vulemp_fem have significance


```{r}
# add female fertility rate, female literacy rate, unemployment rate, female vulnerable employment rate, rate of female in industry, rate of female in service, rate of female in agriculture from the World Development Indicator of the World Bank.

install.packages("WDI")
library(WDI)
newdt <- WDI(indicator = c("SP.DYN.TFRT.IN","SE.ADT.LITR.FE.ZS","SL.UEM.TOTL.FE.ZS", "SL.AGR.EMPL.FE.ZS", "SL.IND.EMPL.FE.ZS","SL.SRV.EMPL.FE.ZS","SL.EMP.VULN.FE.ZS"), start = 1970, end = 2004, extra = T)
names(newdt)[which(names(newdt) == "SP.DYN.TFRT.IN")] <- "fertility"
names(newdt)[which(names(newdt) == "SE.ADT.LITR.FE.ZS")] <- "literacy_fem"
names(newdt)[which(names(newdt) == "SL.UEM.TOTL.FE.ZS")] <- "unemploy_fem"
names(newdt)[which(names(newdt) == "SL.AGR.EMPL.FE.ZS")] <- "agri_fem"
names(newdt)[which(names(newdt) == "SL.IND.EMPL.FE.ZS")] <- "indus_fem"
names(newdt)[which(names(newdt) == "SL.SRV.EMPL.FE.ZS")] <- "serv_fem"
names(newdt)[which(names(newdt) == "SL.EMP.VULN.FE.ZS")] <- "vulemp_fem"

# merge datasets

names(newdt)[which(names(newdt) == "iso3c")] <- "stateid"
newdt1 <- subset(newdt, select = c(stateid, year, fertility, literacy_fem, unemploy_fem, agri_fem, indus_fem, serv_fem, vulemp_fem ))
newdt2 <- newdt1[176:9310,]
as.data.frame(newdt2)
all<- merge(data, newdt2, by.x = c("stateid", "year"), by.y = c("stateid", "year"), all.y=T,all.x=F,sort=T)
as.data.frame(all)

```

```{r}
# rename public spending on education
all <- all %>% 
  rename(spendedc = public_spending_on_education__to)

```


### subset dataset
```{r}
# problem: different datasets cannot compare in likelihood ratio test
# I have to use subset to create a new dataset for specific explanatory variables otherwise I will have 0 obs if I use na.omit for all the variables.
library(dplyr)
all.s <- subset(all, select = c( fertility, vulemp_fem, 
         wecon, lwecon1, lwecon2, lwecon3,
         wosoc, lwosoc1, lwosoc2, lwosoc3, 
         wopol, lwopol1, lwopol2, lwopol3, 
         lsanction_hse, laggdplog,lagsanctiongdp, lagsanctioncost,
         lnsanction_hse_dur, lagsanctdurgdp, lmultisanct, laghumansanction,
         lagpolity2, lagtradelog, lagwar, lagcwar, peace, 
         asia, lamerica, ssafrica, nafrme))

### lag one year for fertility and vulemp_fem
all.s <- all.s %>%
   mutate(lagfert = lag(fertility, n = 1, default = NA))

all.s <- all.s %>%
   mutate(lagvulempfem = lag(vulemp_fem, n = 1, default = NA))

# delete missing values
 all.s <- na.omit(all.s)
```

```{r}
#check correlation for women's economic rights and female vulunrable employment
wecon.num <- as.numeric(all.s$wecon)
cor(wecon.num, all.s$vulemp_fem)
```


### vglm model 1, economic rights

## note: add "lwecon1 + lwecon2 + lwecon3". lwecon0 is reference

### fertility and vulnerable has three-star significant effect on model 1 and model 2.

```{r}
# use lagfert + lagvulempfem
# vglm economic rights

modeco1.vglm <- vglm(wecon ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwecon1 + lwecon2 + lwecon3, data=all.s, 
                 family = cumulative(link= "logitlink", parallel = TRUE))

# new vglm economic rights
modeco1.vglm.new <- vglm(wecon ~ lagfert + lagvulempfem + lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwecon1 + lwecon2 + lwecon3, data=all.s,family = cumulative(link= "logitlink", parallel = TRUE))

## Check parallelism
sink("modpo1vglm.test.txt")
is.parallel(modeco1.vglm.new)
sink

# new tables
library(texreg)
texreg(list(modeco1.vglm, modeco1.vglm.new), include.ci = FALSE)

# likelihood ratio test for differences in models
library(lmtest)
#lrtest(modeco1.vglm, modeco1.vglm.new) ## error

library(nnet) 

sink("lr.eco1.txt")
anova(modeco1.vglm, modeco1.vglm.new, test.statistic="LR", type = 'I', error.estimate=c("pearson", "dispersion", "deviance")) # significance
sink()


## AIC
AIC(modeco1.vglm) # 
AIC(modeco1.vglm.new) #  this one is slightly better
```

### vglm model 2, economic rights

```{r}
# model 2
# add lwosoc1 + lwosoc2 + lwosoc3
# old
modeco2.vglm <- vglm(wecon ~ laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme +lwecon1 + lwecon2 + lwecon3, data=all.s, 
            family = cumulative(link= "logitlink", parallel = TRUE))
# delete " + lwecon1 + lwecon2 + lwecon3"

# new
modeco2.vglm.new <- vglm(wecon ~ lagfert + lagvulempfem + laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwecon1 + lwecon2 + lwecon3, data=all.s, 
                        family = cumulative(link= "logitlink", parallel = TRUE))


#tables
texreg(list(modeco2.vglm, modeco2.vglm.new), include.ci = FALSE)

# likelihood ratio test for differences in models

# lrtest(modeco2.vglm, modeco2.vglm.new) # error

sink("lr.eco2.txt")
anova(modeco2.vglm, modeco2.vglm.new, test.statistic="LR", type = 'I', error.estimate=c("pearson", "dispersion", "deviance")) # significance
sink()

## AIC
AIC(modeco2.vglm) 
AIC(modeco2.vglm.new) # this ACI is smaller
```


## vglm model 1, political rights
### fertility and has three-star signifcant influence and vulunrable has one star on model 1 and model 2
```{r}
# add lwopol1 + lwopol2+  lwopol3
# vglm political rights
modpo1.vglm <- vglm(wopol ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwopol1 + lwopol2+  lwopol3, data=all.s, 
                 family = cumulative(link= "logitlink", parallel = TRUE))

# new vglm political rights
modpo1.vglm.new <- vglm(wopol ~ lagfert + lagvulempfem + lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwopol1 + lwopol2+  lwopol3, data=all.s,family = cumulative(link= "logitlink", parallel = TRUE))

## Check parallelism
is.parallel(modpo1.vglm.new)

# new tables
library(texreg)
texreg(list(modpo1.vglm, modpo1.vglm.new), include.ci = FALSE)

# likelihood ratio test for differences in models
#library(lmtest)
#lrtest(modpo1.vglm, modpo1.vglm.new) # error: 'list' object cannot be coerced to type 'double'

#library(nnet) 
sink("lr.po1.txt")
anova(modpo1.vglm, modpo1.vglm.new, test.statistic="LR", type = 'I', error.estimate=c("pearson", "dispersion", "deviance")) # significance difference
sink()

## AIC
AIC(modpo1.vglm)
AIC(modpo1.vglm.new) # this one is  better

```

### vglm model 2, political rights
```{r, warning=FALSE}
# model 2
# old
modpo2.vglm <- vglm(wopol ~ laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwopol1 + lwopol2+  lwopol3, data=all.s,
            family = cumulative(link= "logitlink", parallel = TRUE))

# new
modpo2.vglm.new <- vglm(wopol ~ lagfert + lagvulempfem + laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwopol1 + lwopol2+  lwopol3, data=all.s, 
                        family = cumulative(link= "logitlink", parallel = TRUE))


#tables
texreg(list(modpo2.vglm, modpo2.vglm.new), include.ci = FALSE)

# likelihood ratio test for differences in models

# lrtest(modpo2.vglm, modpo2.vglm.new) # error

sink("lr.po2.txt")
anova(modpo2.vglm, modpo2.vglm.new, test.statistic="LR", type = 'I', error.estimate=c("pearson", "dispersion", "deviance")) # significance
sink()

## AIC
AIC(modpo2.vglm)
AIC(modpo2.vglm.new) # this one is  better
```

## vglm model 1, social rights
## add lwosoc1 + lwosoc2 + lwosoc3
### fertility and vulunrable has three star
```{r, warning=FALSE}
# vglm model 1, social rights
modso1.vglm <- vglm(wosoc ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwosoc1 + lwosoc2 + lwosoc3, data=all.s, 
                 family = cumulative(link = "logitlink", parallel = TRUE))


# add fertility, literacy_fem, vulemp_fem, serv_fem, indus_fem, agri_fem, vulemp_fem, seceducattained, terteducattained, spendedc 

# new vglm model 1, social rights
modso1.vglm.new <- vglm(wosoc ~ lagfert + lagvulempfem + lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwosoc1 + lwosoc2 + lwosoc3, data=all.s, 
                 family = cumulative(link = "logitlink", parallel = TRUE))

# table
texreg(list(modso1.vglm, modso1.vglm.new), include.ci = FALSE)

# likelihood ratio test for differences in models

## lrtest(modso1.vglm, modso1.vglm.new)
sink("lr.so1.txt")
anova(modso1.vglm, modso1.vglm.new, test.statistic="LR", type = 'I', error.estimate=c("pearson", "dispersion", "deviance")) # significance
sink()

## AIC
AIC(modso1.vglm)
AIC(modso1.vglm.new) # this one is smaller, so it is better
```


### vglm model 2, social rights
```{r}
# old vglm model 2, social rights
modso2.vglm <- vglm(wosoc ~ laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwosoc1 + lwosoc2 + lwosoc3, data=all.s, 
                 family = cumulative(link = "logitlink", parallel = TRUE, reverse = TRUE))

# new  vglm model 2, social rights
modso2.vglm.new <- vglm(wosoc ~ lagfert + lagvulempfem + laggdplog + lnsanction_hse_dur + lagsanctdurgdp + lagsanctioncost + lmultisanct + laghumansanction + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme + lwosoc1 + lwosoc2 + lwosoc3, data=all.s, 
                 family = cumulative(link = "logitlink", parallel = TRUE, reverse = TRUE))

# table
texreg(list(modso2.vglm, modso2.vglm.new), include.ci = FALSE)

# likelihood ratio test for differences in models

# lrtest(modso2.vglm, modso2.vglm.new)
sink("lr.so2.txt")
anova(modso2.vglm, modso2.vglm.new, test.statistic="LR", type = 'I', error.estimate=c("pearson", "dispersion", "deviance")) # significance
sink()

## AIC
AIC(modso2.vglm)
AIC(modso2.vglm.new) # this one is smaller, so it is better

```

```{r}
#calculate AIC and BIC
# economic rights
AIC(modeco1.vglm)
AIC(modeco2.vglm)

AIC(modeco1.vglm.new)
AIC(modeco2.vglm.new)

BIC(modeco1.vglm)
BIC(modeco2.vglm)

BIC(modeco1.vglm.new)
BIC(modeco2.vglm.new)


# social rights
AIC(modso1.vglm)
AIC(modso2.vglm)

AIC(modso1.vglm.new)
AIC(modso2.vglm.new)

BIC(modso1.vglm)
BIC(modso2.vglm)

BIC(modso1.vglm.new)
BIC(modso2.vglm.new)

# political rights
AIC(modpo1.vglm)
AIC(modpo2.vglm)

AIC(modpo1.vglm.new)
AIC(modpo2.vglm.new)

BIC(modpo1.vglm)
BIC(modpo2.vglm)

BIC(modpo1.vglm.new)
BIC(modpo2.vglm.new)
```


## Create comparison tables
## I get opposite results.Sanction is positively rather than negatively related to women's rights and GDP (the wealth of country) is negatively instead of positively related to women's rights. The authors' hypotheses may be not incorrect. But it is problematic to use CIRI human rights to measure women's well-being/status since the results do not make sense.

economic rights
```{r}
# polr, old vglm, new vglm  for economic rights
texreg(list(modeco1,modeco2, 
            modeco1.vglm, modeco2.vglm,
            modeco1.vglm.new, modeco2.vglm.new),
       caption = "Women's Economic Rights Models",
       caption.above = TRUE,
       custom.model.names = c("Ord Logit 1",
                              "Ord Logit 2",
                              "Old VGLM 1",
                              "Old VGLM 2",
                              "New VGLM 1",
                              "New VGLM 2"),
      
         custom.coef.names = c("Economic Sanctions", #1
                             "Sanction*GDP", #2
                             "GDP Per Capita", #3
                             "Sanction Costs", #4
                             "Democracy", #5
                             "Economic Openess", #6
                             "War", #7
                             "Civil War", #8
                             "Peace Years", #9
                             "Asia/Pacific", #10
                             "Latin America/Caribbean", #11
                             "Sub-Saharan Africa", #12
                             "M.East/NA.Africa", #13
                             "Initial Female Labor", #14
                             "Initial Female Labor2", #15
                             "Initial Female Labor3", #16
                             "N1", #17
                             "N2", #18
                             "N3", #19
                             "Sanction Duration", #20
                             "Sanction Duration*GDP", #21
                             "Multilateral Sanction", #22
                             "Humanitarian Sanction", #23
                             "Constant1", #24
                             "Constant2", #25
                             "Constant3", #26
                             "Fertility Rate", #27
                             "Female Vulunrable Employment"), #28
        reorder.coef = c(27:28,
                         1:3,
                         20:21,
                         4,
                         22:23,
                         5:16,
                         24:26,
                         17:19),
        scalebox = .8,
       digits = 3,
       center = TRUE,
       leading.zero = TRUE,
       custom.columns = NULL,
       custom.col.pos = NULL,
       ci.force = FALSE,
        booktabs = TRUE,
       dcolumn = TRUE,
       sideways = FALSE,
       longtable = FALSE,
       use.packages = TRUE,
       table = TRUE,
       no.margin = FALSE,
       fontsize = NULL,
       float.pos = ""
)
```

Political rights
```{r}

texreg(list(modpo1,modpo2, 
            modpo1.vglm, modpo2.vglm,
            modpo1.vglm.new, modpo2.vglm.new),
       caption = "Women's Political Rights Models",
       caption.above = TRUE,
       custom.model.names = c("Ord Logit 1",
                              "Ord Logit 2",
                              "Old VGLM 1",
                              "Old VGLM 2",
                              "New VGLM 1",
                              "New VGLM 2"),
      
         custom.coef.names = c("Economic Sanctions", #1
                             "Sanction*GDP", #2
                             "GDP Per Capita", #3
                             "Sanction Costs", #4
                             "Democracy", #5
                             "Economic Openess", #6
                             "War", #7
                             "Civil War", #8
                             "Peace Years", #9
                             "Asia/Pacific", #10
                             "Latin America/Caribbean", #11
                             "Sub-Saharan Africa", #12
                             "M.East/NA.Africa", #13
                             "Initial Female Labor", #14
                             "Initial Female Labor2", #15
                             "Initial Female Labor3", #16
                             "N1", #17
                             "N2", #18
                             "N3", #19
                             "Sanction Duration", #20
                             "Sanction Duration*GDP", #21
                             "Multilateral Sanction", #22
                             "Humanitarian Sanction", #23
                             "Constant1", #24
                             "Constant2", #25
                             "Constant3", #26
                             "Fertility Rate", #27
                             "Female Vulunrable Employment"), #28
        reorder.coef = c(27:28,
                         1:3,
                         20:21,
                         4,
                         22:23,
                         5:16,
                         24:26,
                         17:19),
        scalebox = .8,
       digits = 3,
       center = TRUE,
       leading.zero = TRUE,
       custom.columns = NULL,
       custom.col.pos = NULL,
       ci.force = FALSE,
        booktabs = TRUE,
       dcolumn = TRUE,
       sideways = FALSE,
       longtable = FALSE,
       use.packages = TRUE,
       table = TRUE,
       no.margin = FALSE,
       fontsize = NULL,
       float.pos = ""
)
        


```

social rights
```{r}
texreg(list(modso1,modso2, 
            modso1.vglm, modso2.vglm,
            modso1.vglm.new, modso2.vglm.new),
       caption = "Women's Social Rights Models",
       caption.above = TRUE,
       custom.model.names = c("Ord Logit 1",
                              "Ord Logit 2",
                              "Old VGLM 1",
                              "Old VGLM 2",
                              "New VGLM 1",
                              "New VGLM 2"),
      
         custom.coef.names = c("Economic Sanctions", #1
                             "Sanction*GDP", #2
                             "GDP Per Capita", #3
                             "Sanction Costs", #4
                             "Democracy", #5
                             "Economic Openess", #6
                             "War", #7
                             "Civil War", #8
                             "Peace Years", #9
                             "Asia/Pacific", #10
                             "Latin America/Caribbean", #11
                             "Sub-Saharan Africa", #12
                             "M.East/NA.Africa", #13
                             "Initial Female Labor", #14
                             "Initial Female Labor2", #15
                             "Initial Female Labor3", #16
                             "N1", #17
                             "N2", #18
                             "N3", #19
                             "Sanction Duration", #20
                             "Sanction Duration*GDP", #21
                             "Multilateral Sanction", #22
                             "Humanitarian Sanction", #23
                             "Constant1", #24
                             "Constant2", #25
                             "Constant3", #26
                             "Fertility Rate", #27
                             "Female Vulunrable Employment"), #28
        reorder.coef = c(27:28,
                         1:3,
                         20:21,
                         4,
                         22:23,
                         5:16,
                         24:26,
                         17:19),
        scalebox = .8,
       digits = 3,
       center = TRUE,
       leading.zero = TRUE,
       custom.columns = NULL,
       custom.col.pos = NULL,
       ci.force = FALSE,
        booktabs = TRUE,
       dcolumn = TRUE,
       sideways = FALSE,
       longtable = FALSE,
       use.packages = TRUE,
       table = TRUE,
       no.margin = FALSE,
       fontsize = NULL,
       float.pos = ""
)
```



## note: I also take into account the dataset problems. I may miss some critical data or delete too many NAs, so that's why I got different results.To solve the possible data issues, I apply authors’ original dataset, the merged dataset, and the selected data of the merging dataset to the same VGLM model for economic rights. I get pretty similar results shown in the Table 6, which are all opposite to authors’ findings. Therefore, it is sufficient to believe that Cooper and Peksen may not have a good selection of models and measurement.


## authors' dataset
```{r}
modeco1.vglm.t1 <- vglm(wecon ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme, data=data, 
                 family = cumulative(link= "logitlink", parallel = TRUE))
```

## merge dataset

```{r}
modeco1.vglm.t2 <- vglm(wecon ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme, data=all, 
                 family = cumulative(link= "logitlink", parallel = TRUE))
```

## selected merge dataset
```{r}
modeco1.vglm.t3 <- vglm(wecon ~ lsanction_hse + lagsanctiongdp + laggdplog + lagsanctioncost + lagpolity2 + lagtradelog + lagwar + lagcwar + peace + asia + lamerica + ssafrica + nafrme, data=all.s, 
                 family = cumulative(link= "logitlink", parallel = TRUE))
```

```{r}
library(texreg)
texreg(list(modeco1.vglm.t1, modeco1.vglm.t2, modeco1.vglm.t3), include.ci = FALSE,
       caption = "different datasets for women's economic rights model 1",
       caption.above = TRUE,
       scalebox = .8,
       digits = 3,
       center = TRUE,
       leading.zero = TRUE,
       custom.columns = NULL,
       custom.col.pos = NULL,
       ci.force = FALSE,
        booktabs = TRUE,
       dcolumn = TRUE,
       sideways = FALSE,
       longtable = FALSE,
       use.packages = TRUE,
       table = TRUE,
       no.margin = FALSE,
       fontsize = NULL,
       float.pos = ""
       )
```


## Model diagnostics

### Working Residuals
```{r}
# economic rights models, working residuals
resid.work.eco1<-resid(modeco1.vglm.new,type="working")
resid.work.eco1[1:10]
resid.work.eco2<-resid(modeco2.vglm.new,type="working")
resid.work.eco2[1:10]

# social rights models, working residuals
resid.work.so1<-resid(modso1.vglm.new,type="working")
resid.work.so1[1:10]
resid.work.so2<-resid(modso2.vglm.new,type="working")
resid.work.so2[1:10]

# political rights models, working residuals
resid.work.po1<-resid(modpo1.vglm.new,type="working")
resid.work.po1[1:10]
resid.work.po2<-resid(modpo2.vglm.new,type="working")
resid.work.po2[1:10]
```

### Pearson residuals
```{r}
# economic rights models, Pearson residuals
resid.pearson.eco1<-resid(modeco1.vglm.new,type="pearson")
resid.pearson.eco1[1:10]

resid.pearson.eco2<-resid(modeco2.vglm.new,type="pearson")
resid.pearson.eco2[1:10]

# social rights models, Pearson residuals
resid.pearson.so1<-resid(modso1.vglm.new,type="pearson")
resid.pearson.so1[1:10]

resid.pearson.so2<-resid(modso2.vglm.new,type="pearson")
resid.pearson.so2[1:10]

# political rights models, Pearson residuals
resid.pearson.po1<-resid(modpo1.vglm.new,type="pearson")
resid.pearson.po1[1:10]

resid.pearson.po2<-resid(modpo2.vglm.new,type="pearson")
resid.pearson.po2[1:10]

```

### Response residuals
```{r}
## note: deviance residuals are nulls. weird... so I use response residuals

# economic rights models, Response residuals
resid.response.eco1<-resid(modeco1.vglm.new,type="response")
resid.response.eco1[1:10]

resid.response.eco2<-resid(modeco2.vglm.new,type="response")
resid.response.eco2[1:10]

# social rights models, Response residuals
resid.response.so1<-resid(modso1.vglm.new,type="response")
resid.response.so1[1:10]

resid.response.so2<-resid(modso2.vglm.new,type="response")
resid.response.so2[1:10]

# political rights models, Response residuals
resid.response.po1<-resid(modpo1.vglm.new,type="response")
resid.response.po1[1:10]

resid.response.po2<-resid(modpo2.vglm.new,type="response")
resid.response.po2[1:10]
```


# plots the Pearson residuals versus the linear predictors (M plots) and plots the Pearson residuals versus the hat values (M plots)
```{r}
# plotvglm 
#Thomas Yee, https://rdrr.io/cran/VGAM/man/plotvglm.html
## plotvglm function plots the Pearson residuals versus the linear predictors (M plots) and plots the Pearson residuals versus the hat values (M plots).

# economic rights models
png(file="eco.resd.png")
par(mfrow=c(1,1))
## model 1
plotvglm(modeco1.vglm.new, which = "(All)")
## model 2
plotvglm(modeco2.vglm.new, which = "(All)")
dev.off()

# social rights models
png(file="so.resd.png")
par(mfrow=c(1,1))
## model 1
plotvglm(modso1.vglm.new, which = "(All)")
## model 2
plotvglm(modso2.vglm.new, which = "(All)")
dev.off()

# political rights models
png(file="po.resd.png")
par(mfrow=c(1,1))
## model 1
plotvglm(modpo1.vglm.new, which = "(All)")
## model 2
plotvglm(modpo2.vglm.new, which = "(All)")
dev.off()


```



